---
title: "Homework1_23220422"
author: Chia Chui Hsien
format: html
editor: visual
execute: 
  echo: false
---

# Gapminder Dataset - Descriptive Statistics and Plot

## 1. Description of dataset

The dataset used in this analysis is **gapminder** which contains data on GDP per capita, population and life expectancies for several countries over many decades.

## 2. Preview of dataset

```{r}
suppressWarnings(library(gapminder))
library(knitr)

kable(head(gapminder), format = "html", caption = "<strong><span style='color:blue'>Preview of Gapminder Dataset</span></strong>")
```

## 3. Descriptive statistics

```{r}
kable(summary(gapminder), format = "html", caption = "<strong><span style='color:blue'>Descriptive Statistics of Gapminder</span></strong>")

```

**Comments**\
The data covers countries across the different continents: Africa, Americas, Asia, Europe and Oceania from 1952 to 2007. For **Life Expectancy**, the median is 60.71 years and the mean is slightly lower at 59.47 years. The maximum is 82.6 years and the minimum is very low at 23.6 years. For **Population**, the median is 7 million people, the mean about 30 million people, the maximum is 1.3 billion people and the minimum is about 60,000 people. For **GDP per capita**, the median is \$3532, the mean, \$7215, the maximum is about \$114,000 and the minimum is \$241. Both **Population** and **GDP per capita** are significantly positively skewed.

## 4. Plot: Comparing life expectancies of selected countries

*The 5 countries (*India, United States, China, Indonesia, Pakistan) *selected for this plot were the top 5 most populated countries in 2023.*

```{r}
suppressWarnings(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))

# Selected countries: top 5 most populated countries in 2023
countries <- c("India", "United States", "China", "Indonesia", "Pakistan")

life_exp <- gapminder |>
  filter(country %in% countries)

# Line plot for 5 countries' life expectancy for period 1952-2007
ggplot(life_exp, aes(x = year, y = lifeExp, color = country)) +
  geom_line() +
  labs(
    title = "Life Expectancy (1952–2007)",
    x = "Year",
    y = "Life Expectancy (Years)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))

```

**Comments**\
All 5 countries experienced an increase in life expectancy from 1952 to 2007, however China's life expectancy dipped drastically between 1957-1962 before rising from 1963 to 2007. Indonesia increased the most significantly (from 37 years in 1952 to 71 years in 2007) while the increase for United States was more moderate (from 68 years to 78 years). It is also noted that United States had a much higher life expectancy (68 years) at the beginning of the period compared to the remaining 4 countries which ranged from 37 to 44 years.
